---
title: "12 Intro to RMarkdown 2"
author: "Alyssa"
date: "2021-03-03"
output:  
  html_document: 
    theme: lumen
    highlight: tango
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

<center><font size="5">
Week 6b  
Intro to RMarkdown 2  
</center></font>

# Outline of class:
Working with RMarkdown today:  
1. Working with figures  
2. Working with tables  
3. Themes  
4. Advanced outputs  
  
Lab:  
1. Practice with RMarkdown  

# Libraries
```{r, message=FALSE, warning=FALSE}
library(palmerpenguins)
library(tidyverse)
library(here)
library(kableExtra)
```


# 1. Working with Figures
aspects of the figure right in our markdown document:  
* change size  
* change resolution  
* change "device" (i.e. is it a .png. .pdf, etc.)  
* save it in a specific folder  

```{r}
penguinsPlot <- penguins %>%
  ggplot(aes(x=species, y=flipper_length_mm, color=island)) +
  geom_boxplot() +
  theme_bw()
penguinsPlot
```

## a. Change size
Some options:

* out.width="70%" (makes it 70% of the width)  
* fig.width=3 (makes the width 3 inches)  
* fig.height=4 (makes the width 4 inches)  
* fig.asp=1.2 (changes the aspect ratio to 1.2)  
* fig.align="center"  
* We will walk through each of these examples together  

[helpful tips](https://r4ds.had.co.nz/graphics-for-communication.html#figure-sizing)  


{r, out.width="70%"}
```{r, out.width="70%", echo=FALSE}
penguinsPlot
```

{r, fig.align='center', fig.asp=1.2}
```{r, fig.align='center', fig.asp=1.2, echo=FALSE}
penguinsPlot
```


## b. Save your figure
ggsave or more automated ways to export your figures in rmarkdown  

set the file path for your figure using:  

* fig.path (you can put this in a global option to send all your figures there)  
* name your code chunk (this will be the name of the figure)  

Place between curly braces  

* {r label}  

Separated options with commas  

* {r label, option1=value}  

Set a figure caption  

* fig.cap="This is a boxplot of penguins  

If you just want the plot and not the code then set echo=FALSE


### figure captions
```{r, fig.align='center', fig.cap="Penguin flipper lengths", echo=FALSE}
penguinsPlot
```


### label code chunks
label is inserted after the r and a space  
{r PenguinFig}
```{r PenguinFig, fig.align='center', fig.cap="Penguin flipper lengths", echo=FALSE}
penguinsPlot
```

### fig.path
add this to chunk header  
fig.path="../Outputs/"   

the name given in the header will be used to name the file  
this can also be added to the global options in the setup   
```{r PenguinFig_fig.path, fig.align='center', fig.cap="Penguin flipper lengths", echo=FALSE, fig.path="../Outputs/"}
penguinsPlot
```


# 2. Tables
## a. type tables by hand  

| Time  | Session | Topic    |
|:------|:-------:|---------:|
| 01:00  | 1       | Anatomy |
| 01:50  |         | *Break* |
| 02:00  | 2       | Tables  |
| 02:45  |         | *Break* |

## b. alignment 
these line here, gives the alignment  
|:------|:-------:|---------:|  
| _left_ | _center_| _right_ |  
the alignment is given by the poition of the : 

## c. kable
install.packages("kableExtra")
library(kableExtra)

make a kable table
```{r}
penguins %>%
  group_by(species) %>%
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl()
```


### themes like alterations
```{r}
penguins %>%
  group_by(species) %>%
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl() %>%
  kable_classic()
```

### row specific changes
```{r}
penguins %>%
  group_by(species) %>%
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl() %>%
  kable_classic() %>%
  row_spec(2, bold = T, color = "white", background = "red")
```

## d. table packages
There are tons of packages that make cool and beautiful tables in rmarkdown.

Check out this blog with lots of examples [here](https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r/)  

# 3. Themes
tons of themes for the whole markdown file  
[theme packages](https://www.datadreaming.org/post/r-markdown-theme-gallery/)

R comes with the bootswatch themes  

package names:  

* prettydoc  
* rmdformats  
* tufte  


# 4. Advanced Outputs

[list of types of outputs](https://rmarkdown.rstudio.com/lesson-9.html)

github_document       allows you to view it nicely on github  

each type of output has different YAML settings


# Help with YAML
Check out the {ymlthis} package which has a GUI for writing yml code.  

`install.packages("ymlthis")`   

Click __Addins__ then scroll to "write new RMarkdown or YAML file"

