---
title: "6b Lab"
author: "Alyssa"
date: "2021-03-03"
output: 
  rmdformats::downcute: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.path="../Outputs/"}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

<center><font size="5"><b>
Week 6b: Lab  
Intro to RMarkdown 2  
</center></font></b>

__Using any data we have used in class:__

* Put in this weeks data folder  
* Create at least one table and one figure in an Rmarkdown file.  
* Create the most beautiful RMarkdown output you can during lab hours in either an html, github doc, or pdf format.  
* Make sure you have headings with clear explanations of what you are doing.  
* Practice using markdown formatting with bold, italics, and lists.  
* Make sure your outputs and scripts are saved in the appropriate folders.  
* Keep proper coding etiquette (figures need to be pub quality as usual).   


# Libraries 
```{r, message=FALSE, warning=FALSE}
library(palmerpenguins)
library(tidyverse)
library(here)
library(kableExtra)
library(rmdformats)
```

# Data
I have chosen to use the palmerpenguins data. I have made a long format of the data set as well as matching set of means. The means are taken of all variables by year, island, species, and sex.
```{r}
penguins <- glimpse(penguins) 

penguins <- penguins %>% relocate(year, island, species, sex) %>% filter(sex != "NA")

penguins_long <- penguins %>%
  pivot_longer(cols = c(5:8), 
               names_to="variables", 
               values_to="values")

means <- penguins %>%  
  pivot_longer(cols = c(5:8), 
               names_to="variables", 
               values_to="values") %>%
  group_by(year, island, species, sex, variables) %>% 
  summarise(means=mean(values, na.rm=T))

means_wide <- means %>% pivot_wider(id_cols = c(1:4), 
                                     names_from="variables", 
                                     values_from="means")
```

# Table
This table shows the mean body mass (g) of all the penguins grouped by species, island, and sex. 
```{r Penguins Mean Body Mass (g)}
penguins %>% 
  group_by(species, island, sex) %>%
  summarise(mean = mean(body_mass_g, na.rm = TRUE)) %>% 
  filter(sex != "NA") %>%
  pivot_wider(id_cols=c(1:2), 
              names_from="sex", 
              values_from="mean") %>% 
  rename("Species" = "species", 
         "Island" = "island", 
         "Female" = "female", 
         "Male" = "male") %>%
  kbl(caption = "<b>Penguins Mean Body Mass (g)</b>") %>% 
  kable_classic() %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F) %>% 
  collapse_rows(columns = 1, valign = "top") %>%
  row_spec(row=0, bold = T, background = "aliceblue")
```

# Plot
Plot all variables and their means for each year. The means are shown with the darker markers.
```{r}
penguins_long$variables <- factor(penguins_long$variables, 
                                  levels = c("bill_depth_mm", "bill_length_mm",
                                             "body_mass_g", "flipper_length_mm"),
                                  labels = c("Bill Depth (mm)", "Bill Length (mm)", 
                                             "Body Mass (g)", "Flipper Length (mm)"))
means$variables <- factor(means$variables, 
                                  levels = c("bill_depth_mm", "bill_length_mm",
                                             "body_mass_g", "flipper_length_mm"),
                                  labels = c("Bill Depth (mm)", "Bill Length (mm)", 
                                             "Body Mass (g)", "Flipper Length (mm)"))

penguins_long %>%
  ggplot(aes(x=year, y=values, color=sex, group=sex, shape=factor(species))) +
  geom_jitter(alpha=0.3, position = position_jitterdodge(0.2)) + 
  geom_point(data=means, aes(x=year, y=means, shape=factor(species))) +
  facet_grid(variables~island, scale="free_y", 
             labeller = label_wrap_gen(multi_line = TRUE, width = 12)) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", 
        legend.box = "vertical", legend.box.just = "left", legend.margin = margin(0),
        legend.key.size = unit(12, "points"), legend.text = element_text(size=7), 
        legend.title = element_text(size=9), legend.spacing.y = unit(0, "lines"), 
        strip.background = element_rect(fill = "white", color = "white"), 
        strip.text = element_text(size=10), 
        strip.text.y = element_text(angle = 0, size=8, hjust = 0, 
                                         margin = margin(2, 0, 0, 1, 
                                                unit = "pt")),
        axis.title = element_text(size=9), axis.text = element_text(size=6)) +
  labs(x = "Bill Depth (mm)", y = "Bill Length (mm)", color="Sex", shape="Species") +
  scale_x_continuous(breaks = seq(from=2007, to=2009, by=1)) 
```


## Bill Depth v Bill Length
plot of the bill length by the depth. the means are included and shown by the darker markers and labeled with their year. 
```{r}
penguins %>% filter(sex != "NA") %>% 
  ggplot(aes(x=bill_depth_mm, y=bill_length_mm, color=sex, shape=factor(species))) +
  geom_point(alpha=0.3) +
  geom_point(data=means_wide, alpha=1) +
  geom_text(data=means_wide, aes(label=year), size=2, nudge_x = 0.25, nudge_y = 0.25,
            check_overlap = T) +
  facet_wrap(island~.) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", 
        legend.box = "vertical", legend.box.just = "left",
        legend.text = element_text(size=8), legend.title = element_text(size=9), 
        legend.margin = margin(0), legend.spacing.y = unit(0, "lines"), 
        strip.background = element_rect(fill = "white", color = "white"), 
        strip.text = element_text(size=10), axis.title = element_text(size=9)) +
  labs(x = "Bill Depth (mm)", y = "Bill Length (mm)", color="Sex", shape="Species") + 
  ggsave(here("Week_6", "Outputs", "Bill_Depth_Length.png"))
```


## flipper length v. body mass
plot of the body mass by the flipper length. the means are included and shown by the darker markers and labeled with their year.
```{r}
penguins %>% filter(sex != "NA") %>% 
  ggplot(aes(x=flipper_length_mm, y=body_mass_g, color=sex, shape=factor(species))) +
  geom_point(alpha=0.3) +
  geom_point(data=means_wide, alpha=1) +
  geom_text(data=means_wide, aes(label=year), size=2, nudge_x = 4, nudge_y = -0.7,
            check_overlap = T) +
  facet_wrap(island~.) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", 
        legend.box = "vertical", legend.box.just = "left", 
        legend.box.margin = margin(0, 0, 0, 0), legend.key.size = unit(12, "points"),
        legend.text = element_text(size=8), legend.title = element_text(size=9), 
        legend.margin = margin(0), legend.spacing.y = unit(0, "lines"), 
        strip.background = element_rect(fill = "white", color = "white"), 
        strip.text = element_text(size=10), axis.title = element_text(size=9), 
        axis.text = element_text(size=6)) +
  labs(x = "Flipper Length (mm)", y = "Body Mass (g)", color="Sex", shape="Species")
```

