---
title: "15_Advanced_Plotting"
author: "Alyssa"
date: "3/22/2021"
output: 
  html_document: 
    theme: spacelab
    highlight: tango
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

<center>

<font size="5">

Week 8a\
Advanced Plotting

</center>

</font>

# Outline of class

Advanced plotting

1.  {patchwork} - bringing plots together
2.  {ggrepel} - repel your labels on your plots
3.  {gganimate} - add some animations to your plots
4.  {magick} - photo processing

I am just going to give you a little taste of what each of these packages can do. Please check out the documentation for each to delve deeper.

Lab

1.  Good plot/Bad plot lab presentations!

# Libraries

## install

```{r, eval=FALSE}
install.packages("patchwork") # for bringing plots together
install.packages("ggrepel") # for repelling labels
install.packages("gganimate") # animations
install.packages("magick") # for images
```


## load
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(patchwork)
library(ggrepel)
library(gganimate)
library(magick)
library(palmerpenguins)
```


# 1. Patchwork
Easily bring your plots together  

Let's start with two simple plots from the Palmer penguin dataset.  

Plot 1:
```{r, message=FALSE, warning=FALSE, out.width='60%', fig.align='center'}
p1<-penguins %>%
  ggplot(aes(x = body_mass_g, 
             y = bill_length_mm, 
             color = species))+
  geom_point()

p1
```

Plot 2:
```{r, message=FALSE, warning=FALSE, out.width='60%', fig.align='center'}
p2<-penguins %>%
  ggplot(aes(x = sex, 
             y = body_mass_g, 
             color = species))+
  geom_jitter(width = 0.2)

p2
```

## Both together using patchwork
```{r, message=FALSE, warning=FALSE, fig.asp=.5, fig.align='center', out.width='80%'}
p1 + p2
```

## Group the legends together
```{r, message=FALSE, warning=FALSE, fig.asp=.5, fig.align='center', out.width='80%'}
p1+p2 +
  plot_layout(guides = 'collect')
```

## Add Labels
plot_annotation
```{r, message=FALSE, warning=FALSE, fig.asp=.5, fig.align='center', out.width='80%'}
p1+p2 +
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = 'A')
```

## vertical layot
use / instead of +
```{r plot1, message=FALSE, warning=FALSE, fig.asp=.5, fig.align='center', out.width='80%', eval=FALSE}
p1/p2 + #<<
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = 'A')
```

So many cool ways to bring together and modify plots.  For more info see the many [vignettes here](https://patchwork.data-imaginist.com/articles/patchwork.html)  


# 2. ggrepel
Easy and clear labels for plots  

Use the mtcars dataset that comes with it.  It is data on cars.
```{r, eval = FALSE, warning=FALSE, message=FALSE}
View(mtcars)
```

using regular 'geom_text' causes a lot of overlap 
```{r plot2, eval=FALSE}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) +
  geom_text() + # creates a text label
  geom_point(color = 'red') 
```

## geom_text_repel
Repel the labels
```{r plot3, eval=FALSE}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) +
  geom_text_repel() +
  geom_point(color = 'red') 
```

## geom_label_repel
repel labels  
labels are put in a box
```{r plot4, eval=FALSE}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) +
  geom_label_repel() + # repel them
  geom_point(color = 'red') 
```

For more cool things you can do with [ggrepel see here](https://ggrepel.slowkow.com/articles/examples.html)


# gganimate
Make your figure an animation!  
\
Let's go back to our penguin plot, but animate the figure by year.  
Our static plot.
```{r plot5, warning=FALSE, message=FALSE, eval = FALSE}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() 
```

## Add a transition

'transition_states'
  year                    what are we animating by  
  transition_length       The relative length of the transition  
  state_length            The length of the pause between transitions  
\
Here see one year at a time:
```{r, warning=FALSE, message=FALSE}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(year, 
                    transition_length = 2, 
                    state_length = 1)
```


## Change the ease aesthetics
A few examples:

- linear (default)
- bounce
- sine
- circular
- many more...

```{r, warning=FALSE, message=FALSE}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(year, 
                    transition_length = 2, 
                    state_length = 1) +
  ease_aes("bounce-in-out")
```


## Add a transition title
ggtitle
```{r, warning=FALSE, message=FALSE}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(year, 
                    transition_length = 2, 
                    state_length = 1) +
  ease_aes("bounce-in-out") +
  ggtitle('Year: {closest_state}')
```


## Save it as a .gif
anim_save
```{r, warning=FALSE, message=FALSE}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(year, 
                    transition_length = 2, 
                    state_length = 1) +
  ease_aes("bounce-in-out") +
  ggtitle('Year: {closest_state}') +
  anim_save(here("Week_8","Outputs","mypengiungif.gif"))
```

For more things to do [see here](https://gganimate.com/articles/gganimate.html)


# 3. magick
Advanced image processing

## Put a penguin in it
Read in an image of a pengiun (can be on your computer or the internet)
```{r}
penguin<-image_read("https://pngimg.com/uploads/penguin/pinguin_PNG9.png")

penguin
```

To put it on a plot you first need to save your plot as an image
```{r, warning = FALSE, message=FALSE,out.width='50%', fig.align='center'}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  ggsave(here("Week_8","Outputs","penguinplot.png"))
```

Read in the plot as a magick image and make a composite plot  
  offset uses "+X+Y"
```{r, warning = FALSE, message=FALSE, fig.align='center', out.width='40%'}
penplot<-image_read(here("Week_8","Outputs","penguinplot.png"))
out <- image_composite(penplot, penguin, offset = "+70+30")
out
```

You can do this with gifs too
  gravity uses "center", "top", etc.
```{r, warning=FALSE, message=FALSE, out.width='40%', fig.align='center'}
pengif<-image_read("https://media3.giphy.com/media/H4uE6w9G1uK4M/giphy.gif")
outgif <- image_composite(penplot, pengif, gravity = "center")
animation <- image_animate(outgif, fps = 10, optimize = TRUE)
animation
```


For more magick plotting [see here](https://cran.r-project.org/web/packages/magick/vignettes/intro.html)


