---
title: "18_Getting_Help"
author: "Alyssa"
date: "4/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Outline of Class

Getting Help with R

1. Troubleshooting

2. Common Mistakes

3. Reproducible examples {reprex}

4. Copying and pasting {datapasta}

Lab

1. Make a Reprex

# Libraries
Make a reproducible example to ask for help
```{r, eval = FALSE}
install.packages('reprex') #reproducible example
install.packages('datapasta') #copy and paste
install.packages('styler') # copy and paste in style
```

# Making your first reprex

Let's say you are trying to make a plot with the mpg dataset and you hit an error.
```{r, error=TRUE}
library(tidyverse)

mpg %>%
  ggplot(aes(x = displ, y = hwy))%>%
  geom_point(aes(color = class))
```


**Go to Addins -> Render Reprex -> Paste to the program you want.  
Make sure to "append session info"**


# Using real data

Often you are working with real data and run into an issue.  You need the example to be **reproducible** and simple.  Other people on the internet do not have your data so what should we do?  

Open the *stars.csv* file and copy and paste the first few lines into R.


# datapasta

{datapasta} helps you copy and past snippets of your data in a reproducible way.  

This time copy the code and got to Addins -> paste as tribble and then name it something.  

```{r}
data<-tibble::tribble(
    ~lat,    ~long, ~star_no,
  33.548, -117.805,      10L,
  35.534, -121.083,       1L,
  39.503, -123.743,      25L,
  32.863,  -117.24,      22L,
   33.46, -117.671,       8L,
  33.548, -117.805,       3L
  )


data
```

You can now use this with {reprex} to make a reproducible example that can fully run with your own data and code. 

# Do's and don't of reprex

1. Use the smallest, simplest, most built-in data possible.
  - e.g. using datapasta  
  
2. Include commands on a strict “need to run” basis.  
  - Ruthlessly strip out anything unrelated to the specific matter at hand.  
  - Include every single command that is required, e.g. loading specific packages  
  
3. Pack it in, pack it out, and don’t take liberties with other people’s computers. You are asking people to run this code!  
  - Don’t start with rm(list = ls()). It is anti-social to clobber other people’s workspaces.  
  - Don’t start with setwd("C:\Users\jenny\path\that\only\I\have"), because it won’t work on anyone else’s computer.  
  - Don’t mask built-in functions, i.e. don’t define a new function named c or mean.  
  - If you change options, store original values at the start, do your thing, then restore them.  
  - If you create files, delete them when you’re done  
  
4. Use good coding style 



The reprex code:  
- Must run and, therefore, should be run by the person posting. No faking it.  
- Should be easy for others to digest, so they don’t necessarily have to run it. You are encouraged to include selected bits of output.  
- Should be easy for others to copy + paste + run, if and only if they so choose. Don’t let inclusion of output break executability.  

.foot-note[https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html]
