---
title: "13 Intro to Maps pt1"
author: "Alyssa"
date: "2021-03-08"
output: 
  html_document: 
    theme: lumen
    highlight: tango
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

<center><font size="5">
Week 7a  
Intro to Maps part 1  
</center></font>

# Outline of Class
Working with Maps:  

1. Highlight some various spatial packages  
2. Simple mapping using the {maps} package with ggplot  

Lab:  

1. Make a map  


# New text editor
download new Rstudio 
shows up on top right as A
{r}
1+1

# Libraries
```{r, eval = FALSE}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
```

# Data
population in California by county  
number of seastars at different field sites  
```{r}
popdata<-read_csv(here("Week_7","Data","CApopdata.csv"))
stars<-read_csv(here("Week_7","Data","stars.csv"))
```



# 1. Highlight some various spatial packages

## Map Packages
Today we will use these two:
```{r, message=FALSE, warning=FALSE}
install.packages(c("maps", "mapproj"))
install.packages("mapdata")
```

Other very helpful packages:  

* __sf__    (This is for the serious GIS people. If you want to do actual spatial analyses this is the package to use)  
* __ggmap__   (Uses the google maps API to make maps. Really easy to use, but requires a credit card for the API...)  
* __tmap__    (static and interactive maps)  
* __leaflet__   (interactive html maps)  

Helpful references:  

1. The required reading for today  
2. Advanced maps  
3. Bathymetry example  
4. Earth Lab  


## {maps} package
The {maps} package is a combination of functions that pair well with ggplot and base layers for maps (i.e. polygons of regions of interest).  

map_data("location") is the function used to pull out whatever base layer that you want.  

### Extract data from mapdata package
Data for the entire world:
```{r}
world<-map_data("world")
head(world)
```


Polygons for different countries:
```{r}
usa<-map_data("usa")
head(usa)

italy<-map_data("italy")
head(italy)
```


### Extract data and different resolutions.  

For example for data that also has polygons for each state in the US use:
```{r}
states<-map_data("state")
head(states)
```


For example for data that also has polygons for each county in the US use:  
```{r}
counties<-map_data("county")
head(counties)
```


### Structure of the data

* **long** is longitude. Things to the west of the prime meridian are negative.  

* **lat** is latitude  

* **order** This just shows in which order ggplot should “connect the dots”  

* **region and subregion** tell what region or subregion a set of points surrounds.  

* **group**  
  + This is very important!  
  + ggplot2’s functions can take a group argument which controls (amongst other things) whether adjacent points should be connected by lines.  
    + If they are in the same group, then they get connected, but if they are in different groups then they don’t.  
    + Essentially, having to points in different groups means that ggplot “lifts the pen” when going between them.  
    
# 2. Simple mapping using the {maps} package with ggplot

## Map of the world
```{r}
ggplot()+
  geom_polygon(data = world, aes(x = long, y = lat, group = group))
```

this is what happens when you forget __group__ = __group__ :
```{r}
ggplot()+
  geom_polygon(data = world, aes(x = long, y = lat))
```


Add color to the lines and fill:
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, y = lat, group = group),
               color = "black")
```


Color by country:
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group, 
                   fill = region),
               color = "black") +
  guides(fill = FALSE)
```


## Change the visuals

### Theme
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat,
                   group = group,
                   fill = region),
               color = "black")+
  guides(fill = FALSE)+
  theme_minimal()
```


### Make ocean blue
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group,
                   fill = region),
               color = "black")+
  theme_minimal()+
  guides(fill = FALSE)+
  theme(panel.background = element_rect(fill = "lightblue"))
```


### Change map projection
The earth is NOT flat, but we are attempting to visualize this map in 2D. There are a lot of different projections that you can use. See here for the different projections included in the {mapproj} package

Use a mercator projection
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group,
                   fill = region),
               color = "black")+
  theme_minimal()+
  guides(fill = FALSE)+
  theme(panel.background = element_rect(fill = "lightblue"))+
  coord_map(projection = "mercator",
            xlim = c(-180,180))
```



Use a sinusoidal projection.
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long,
                   y = lat,
                   group = group, 
                   fill = region),
               color = "black")+
  theme_minimal()+
  guides(fill = FALSE)+
  theme(panel.background = element_rect(fill = "lightblue"))+
  coord_map(projection = "sinusoidal",
            xlim = c(-180,180))
```


## Map of just California
Use the states dataset
```{r}
CA_data<-states %>%
  filter(region == "california")
```


### T,P,S: make a map of California
```{r}
ggplot()+
geom_polygon(data = CA_data, 
             aes(x = long, 
                 y = lat, 
                 group = group), 
             color = "black")+
  coord_map()+
  theme_void()
```


## Adding multiple layers of data
Plot the population of every county in California.

First, Get the county level data and join it with our dataset on population  
```{r}
head(counties)[1:3,]
head(popdata)
```

Wrangle the data. Make the column names the same to join the two datasets.  
Join by "subregion"
```{r}
CApop_county<-popdata %>%
  select("subregion" = County, Population)  %>% # rename the county col
  inner_join(counties) %>%
  filter(region == "california")

head(CApop_county)
```

### Map of CA population by county
```{r}
ggplot()+
geom_polygon(data = CApop_county, 
             aes(x = long, 
                 y = lat, 
                 group = group,
                 fill = Population),
             color = "black")+
  coord_map()+
  theme_void()
```


Log scale for easier interpretation
```{r}
ggplot()+
geom_polygon(data = CApop_county, 
             aes(x = long, 
                 y = lat, 
                 group = group,
                 fill = Population),  
             color = "black")+
  coord_map()+
  theme_void() +
  scale_fill_gradient(trans = "log10")
```


### Add a layer of points
Look at the stars dataset. This has the number of seastars per/m2 at different sites in California.
```{r}
head(stars)
```

Add a geom_point layer
```{r}
ggplot()+
geom_polygon(data = CApop_county, 
             aes(x = long, 
                 y = lat, 
                 group = group,
                 fill = Population),  
             color = "black")+
  geom_point(data = stars, # add a point at all my sites
             aes(x = long,
                 y = lat))+
  coord_map()+
  theme_void() +
  scale_fill_gradient(trans = "log10")
```


Make points proportional to number of stars 
```{r}
ggplot()+
geom_polygon(data = CApop_county, 
             aes(x = long, 
                 y = lat, 
                 group = group,
                 fill = Population),  
             color = "black")+
  geom_point(data = stars, # add a point at all my sites 
             aes(x = long, 
                 y = lat,
                 size = star_no))+
  coord_map()+
  theme_void() +
  scale_fill_gradient(trans = "log10")
```

Make a better legend label
```{r}
ggplot()+
geom_polygon(data = CApop_county, 
             aes(x = long, 
                 y = lat, 
                 group = group,
                 fill = Population),  
             color = "black")+
  geom_point(data = stars, # add a point at all my sites 
             aes(x = long, 
                 y = lat,
                 size = star_no))+ 
  coord_map()+
  theme_void() +
  scale_fill_gradient(trans = "log10")+
  labs(size = "# stars/m2")
```


Save it
```{r}
ggplot()+
geom_polygon(data = CApop_county, 
             aes(x = long, 
                 y = lat, 
                 group = group,
                 fill = Population),  
             color = "black")+
  geom_point(data = stars, # add a point at all my sites 
             aes(x = long, 
                 y = lat,
                 size = star_no))+ 
  coord_map()+
  theme_void() +
  scale_fill_gradient(trans = "log10")+
  labs(size = "# stars/m2") +
  ggsave(here("Week_7","Outputs","CApop.pdf"))
```









