---
title: "R Notebook"
output: 
  html_notebook: 
    theme: spacelab
editor_options: 
  markdown:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

<center>

<font size="5">

Week 7b\
Intro to Maps part 2

</center>

</font>

# Outline of Class

Working with Maps part 2:

1.  {ggmaps}

Lab:

1.  Good plot/Bad plot lab

[ggmap article](http://stat405.had.co.nz/ggmap.pdf)\
[Making plots with ggmap](https://www.littlemissdata.com/blog/maps)\
[Setting up ggmap](https://www.findingyourway.io/blog/2018/12/05/2018-12-05-using-ggmap-after-july-2018/)

# Libraries

install ggmaps and ggsn

```{r, eval=FALSE}
install.packages("ggmap")
install.packages("ggsn")
```

load libraries for today

```{r, warning=TRUE, message=TRUE}
library(ggmap)
library(tidyverse)
library(here)
library(ggsn)
```

use your own API IMPORTANT don't put your key to github!! I usually save mine in a text file and import it each time

```{r, eval=FALSE}
register_google(key = "AIzaSyCTLtIVH1tWKFwskcADuO7hlzjZ9wiSVjY")
```

# Data

```{r}
ChemData<-read_csv(here("Week_7","Data","chemicaldata_maunalua.csv"))
glimpse(ChemData)
```

# Get base maps from ggmap

*get_map()* is the function to get a basemap from Google Maps. At it's simplest form, you only need to put in a location.

## Location as a string

```{r, warning=FALSE, message=FALSE}
Oahu <- get_map("Oahu")
```

## Use ggmap to plot the base layer

```{r, out.width="60%", fig.align="center"}
ggmap(Oahu)
```

## Get coordinates for Wailupe on Oahu:

-   Make a data frame of lon and lat coordinates

    -   coordinates for Wailupe

-   Get base layer

-   Plot it

```{r, warning=FALSE, message=FALSE, out.width="60%", fig.align='center'}
WP<-data.frame(lon = -157.7621, lat = 21.27427) 

Map1<-get_map(WP)

ggmap(Map1)
```

## Zoom in on a location

The *zoom* argument, an integer from **3 to 20** specifying how large the spatial extent should be around the center, with 3 being the **continent level** and 20 being roughly the **single building level**.

```{r, warning=FALSE, message=FALSE, out.width="60%", fig.align='center'}
Map1<-get_map(WP,zoom = 17)

ggmap(Map1)
```

## Change the map type

There are lots of different ones. look at ?get_map

Satellite Maps:

```{r, warning=FALSE, message=FALSE, out.width="60%", fig.align='center'}
Map1<-get_map(WP,zoom = 17, maptype = "satellite")

ggmap(Map1)
```

Water Color:

```{r, warning=FALSE, message=FALSE, out.width="60%", fig.align='center'}
Map1<-get_map(WP,zoom = 17, maptype = "watercolor")

ggmap(Map1)
```

# Use the ggmap base layer in any ggplot

```{r, warning=FALSE, message=FALSE, out.width="60%", fig.align='center'}
Map1<-get_map(WP,zoom = 17, maptype = "satellite") 

ggmap(Map1)+
  geom_point(data = ChemData, 
             aes(x = Long, y = Lat, color = Salinity),
             size = 2) + 
  scale_color_viridis_c()
```

# Add scale bar

```{r, warning=FALSE, message=FALSE, out.width="60%", fig.align='center'}
ggmap(Map1)+
  geom_point(data = ChemData, 
             aes(x = Long, y = Lat, color = Salinity), 
             size = 2) + 
  scale_color_viridis_c()+
  scalebar( x.min = -157.766, x.max = -157.758,
           y.min = 21.2715, y.max = 21.2785,
            dist = 250, dist_unit = "m", model = "WGS84", 
            transform = TRUE, st.color = "white",
            box.fill = c("yellow", "white")) 
```

# geocode()

This is what you use if you don't know exact lat and long.\
Use *geocode()* to get exact locations that you can then use in the maps.

```{r, message=FALSE, warning=FALSE}
geocode("the white house")

geocode("California State University, Northridge")
```
