---
title: "7a Lab"
author: "Alyssa"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE, fig.path="../Outputs/"}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```


# Lab Outline
You need to make a map using RMarkdown with one of the following datasets from past Tidy Tuesdays. Make sure, as usual, that you scripts and output are saved appropriately.

Plastic pollution

Trees of San Fransisco

Meteorites

# Libraries
```{r, eval = FALSE}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
library(ggplot2)
library(viridis)
```


# Data
```{r}
sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')
```

separate species name and common name
```{r, eval=FALSE}
sf_trees <- sf_trees %>%
  separate(col = species, 
           into = c("Species", "Common_Name"),
           sep = "::") %>% 
  filter(Species != "Tree(s) ") %>% 
  rename("long" = "longitude", "lat" = "latitude") %>%
  str_trim(sf_trees$Species, side = "right")

landmark_trees <- sf_trees %>% filter(legal_status =="Landmark tree")
```


# Map of California
Get the map data for san francisco county
```{r, eval=FALSE}
map <- map_data("county") %>% filter(region == "california")
  
SF_map <- map_data("county") %>% filter(subregion == "san francisco")

map2 <- map_data("county") %>% filter(subregion %in% c("san francisco", "marin", "san mateo", "contra costa", "alameda"))

landmark_trees$dbh <- as.numeric(landmark_trees$dbh)
landmark_trees %>% group_by(dbh) %>% summarize(n=n())

landmark_trees <- landmark_trees %>% filter(dbh<=152)
```

San Francisco Map Bounds
|       |Lat        | Long     |
|-------|-----------|----------|
|Max    |37.81522   |-122.3208 |  
|Min    |37.70062   |-122.4812 |  


Landmark Trees Map
```{r}
ggplot()+
  geom_polygon(data = SF_map, aes(x = long, y = lat, group = group), 
             fill = "grey", alpha=0.3) +
  geom_point(data=landmark_trees, aes(x = long, y = lat, color=Species), 
             size=1) +
  coord_map(xlim = c(-122.5, -122.32), ylim = c(37.71, 37.81)) + 
  theme_void() +
  theme(legend.position = "none",
        plot.margin = unit(c(0.2,1,0.2,0.1), "cm")) + 
  labs(color="Species", title = "Landmark Trees in San Francisco")
```


# Significant Trees
Get list of the top 6 most common tree species
```{r}
sf_trees %>% filter(legal_status == "Significant Tree") %>% 
  select(tree_id, Species, long, lat) %>% 
  group_by(Species) %>% summarise(n=n()) %>% top_n(6)
```

Plot
```{r}
sf_trees %>% 
  filter(legal_status == "Significant Tree", 
         Species %in% c("Acacia melanoxylon", "Cupressus macrocarpa", 
                        "Myoporum laetum", "Pinus radiata", 
                        "Pittosporum undulatum", "Platanus x hispanica")) %>%
  ggplot()+
  geom_polygon(data = SF_map, aes(x = long, y = lat, group = group), 
             fill = "grey", alpha=0.35, color="black", size=0.4) +
  geom_point(aes(x = long, y = lat, color=Species), size=1, alpha=0.8) +
  coord_map() + 
  theme_void() +
  theme(legend.position = "right", legend.direction = "vertical", 
        legend.title = element_text(size=9), legend.text = element_text(size=8),
        legend.margin = margin(0,0,0,0, "pt"), 
        legend.key.size = unit(1.7, "lines"),
        plot.margin = unit(c(0.2,1,0.2,0.1), "cm")) + 
  labs(title = "Top 6 Significant Tree Species in San Francisco") +
  scale_color_discrete(labels = function(x) str_wrap(x, width = 11)) +
  guides(color = guide_legend(override.aes = list(size=2))) +
  ggsave(here("Week_7", "Outputs", "SF_Trees.png"))
```



