---
title: "7b Lab"
author: "Alyssa Cohen"
date: "2021-03-10"
output: 
  html_document: 
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Good Plot v Bad Plot {.tabset .tabset-fade .tabset-pills}

## Home

```{r, out.width=c('50%', '50%'), fig.show='hold'}

```

::: {.columns}
::: {.column width="50%" data-latex="{0.05\\textwidth}"}
Text describing Good Plot.
:::

::: {.column width="50%" data-latex="{0.05\\textwidth}"}
Text describing bad plot
:::
:::

## Info

<center>

<font size="6">

<p style="color:black">

Good Plot Bad Plot Lab\

</center>

</font>

</p>

[From Dr. Drew Steen]

Assignment given by Dr. Nyssa Silbiger, CSUN Spring 2021 for her Computer Modeling course.

**Due March 22**

The homework is to create two version of similar plots, using any dataset you choose. You must create the plots using ggplot2, post-processing with Adobe Illustrator or similar is not allowed. **This homework is a contest, and the winner will win a prize to be disclosed.**

Submission Details

-   Post a markdown document containing

    -   The plots
    -   The code you used to create them
    -   Your discussion of the plot within the markdown doc to github
    -   Data also needs to be submitted

-   Bad Plot

    -   Should be as bad as possible

        -   Your reading from Claus Wilke's and Healy's textbook describe the many ways that a plot cab become misleading, difficult to interpret, or or just plain ugly, and you should use all of the tools that you can.
        -   Please try to make me, a person who cares a great deal about high-quality data visualization, cry.

    -   You should explain in text all of the ways in which the bad plot is bad

        -   Bullet points are fine

        -   Clearly articulate each principle of data visualization that you have broken

-   Good Plot

    -   Based on the same or similar data to the first plot

    -   Should be good

    -   Write briefly about why this is a successful plot

-   Both Plots

    -   Must be made in ggplot2

    -   Code visible in the Rmarkdown document

    -   The data must be visible

    -   The good plot and the bad plot should be based on the same, or similar, data

    -   Nyssa will post the results on **Twitter**. Please notify me if you would prefer not to have your plots posted publicly. That is fine, but I need to know

## Libraries

```{r, warning=TRUE, message=TRUE}
library(tidyverse)
library(here)
library(extrafont)
library(yarrr)
```

Loaded my computer's fonts with the extrafonts package

I imported the fonts then viewed which loaded.

```{r, eval=FALSE}
font_import()
fonts()
fonttable()
fonttable()[40:45,]
```

## Data

### Metadata

#### Site Characteristics Data

+--------------------+-----------------------------------------------------------+------------------------------+
| Variable           | Description                                               | Values/Units                 |
+====================+===========================================================+==============================+
| name               | Name of the site                                          | character                    |
+--------------------+-----------------------------------------------------------+------------------------------+
| site.letter        | Code associated with each site                            |                              |
+--------------------+-----------------------------------------------------------+------------------------------+
| site.block         | Experimental block (two sites were done at the same time) |                              |
+--------------------+-----------------------------------------------------------+------------------------------+
| parameter.measured | Environmental parameter measured                          | Light (proton flux density), |
|                    |                                                           |                              |
|                    |                                                           | temp (degrees C),            |
|                    |                                                           |                              |
|                    |                                                           | NH4 (ammonium umol/L)        |
+--------------------+-----------------------------------------------------------+------------------------------+
| values             | value of the measurement taken                            |                              |
+--------------------+-----------------------------------------------------------+------------------------------+

Parameters

+------------------------+--------------------------------+----------------------+
| Code                   | Description                    | Unit                 |
+========================+================================+======================+
| Light                  | proton flux density            |                      |
+------------------------+--------------------------------+----------------------+
| temp                   | Temperature                    | C                    |
+------------------------+--------------------------------+----------------------+
| NH4                    | ammonium                       | umol/L               |
+------------------------+--------------------------------+----------------------+
| N.N                    | Nitrate + nitrite              | umol/l               |
+------------------------+--------------------------------+----------------------+
| P                      | phosphate                      | umol/L               |
+------------------------+--------------------------------+----------------------+
| DIN_DIP                | nitrogen to phosphorus ratio   | ratio                |
+------------------------+--------------------------------+----------------------+
| N                      | tissue N in Sargassum          | \%                   |
+------------------------+--------------------------------+----------------------+
| trap.accumulation.rate | sedimentation rate             |                      |
+------------------------+--------------------------------+----------------------+
| CCA.cover              | crustose coralline algae cover | \%                   |
+------------------------+--------------------------------+----------------------+
| algal.cover            | fleshy algae cover             | \%                   |
+------------------------+--------------------------------+----------------------+
| p.acuta.cover          | *Pocillopora acuta* cover      | \%                   |
+------------------------+--------------------------------+----------------------+
| coral.cover            | coral cover                    | \%                   |
+------------------------+--------------------------------+----------------------+
| substrate.cover        | other substrate cover          | \%                   |
+------------------------+--------------------------------+----------------------+

#### Topt Data

+-------------+--------------------------------------------------------------------------+----------------------------+
| Variable    | Description                                                              | Values/Units               |
+=============+==========================================================================+============================+
| fragment.ID | Unique ID of every coral fragment (*Pocillopora acuta*)                  | PA\#\#\_X                  |
+-------------+--------------------------------------------------------------------------+----------------------------+
| rate.type   | The type of metabolic rate measured                                      | R = respiration,           |
|             |                                                                          |                            |
|             |                                                                          | GP = Gross Photosynthesis, |
|             |                                                                          |                            |
|             |                                                                          | C = calcification          |
+-------------+--------------------------------------------------------------------------+----------------------------+
| E           | Activation energy extracted from thermal performance curve               | eV                         |
+-------------+--------------------------------------------------------------------------+----------------------------+
| Eh          | Deactivation energy extracted from thermal performance curve             | eV                         |
+-------------+--------------------------------------------------------------------------+----------------------------+
| lnc         | Rate at a constant temperature extracted from thermal performance curves | umol cm-2 hr-1             |
+-------------+--------------------------------------------------------------------------+----------------------------+
| Th          | Temperature in Kelvin where half the enzymes are inactivated             | K                          |
+-------------+--------------------------------------------------------------------------+----------------------------+
| Topt        | Thermal optima extracted from ther thermal performance curve             | degree C                   |
+-------------+--------------------------------------------------------------------------+----------------------------+
| site.letter | Code associated with each site                                           |                            |
+-------------+--------------------------------------------------------------------------+----------------------------+

```{r}
sites <- read_csv(here("Week_5", "Data", "site.characteristics.data.csv"))

topt <- read_csv(here("Week_5", "Data", "Topt_data.csv"))
```

## Plots

```{r}

  
```

Coral Topt Rates Plot

```{r}
coral_plot <- sites %>% select(name, site.letter) %>% 
  right_join(topt) %>% 
  select(fragment.ID, rate.type, "Activation_Energy"=E, "Rate"=lnc, "Site"= name)

coral_plot_labels <- data.frame(rate.type = unique(coral_plot$rate.type)) %>%
  mutate(Type = recode(rate.type, 
                       "R" ="Respiration",
                       "GP"="Gross Photosynthesis",
                       "C" ="Calcification"))
coral_plot_labels$Activation_Energy <- c(1, 0.28, 0.17)
coral_plot_labels$Rate <- c(1.5, 0.65,0.35)
```

```{r}
coral_plot %>%
  ggplot(aes(x=Activation_Energy, y=Rate, color=Site)) +
  geom_point() +
  facet_wrap(rate.type~., nrow=3, ncol=1, scales = "free",
             labeller=labeller(rate.type=c("R" ="Respiration", 
                                           "GP"="Gross Photosynthesis", 
                                           "C" ="Calcification"))) +
  scale_color_manual(values=unname(piratepal(palette = "pony", length.out = 6))) +
  theme_bw() +
  theme(legend.position = "bottom", plot.background = element_rect(fill=NULL),
        strip.text = element_text(face="bold"))
```

```{r}
coral_plot %>%
  ggplot(aes(x=Activation_Energy, y=Rate, color=rate.type)) +
  geom_point() +
  scale_color_manual(values=unname(piratepal(palette = "pony", length.out = 6))) +
  theme_bw() +
  scale_color_discrete(name = "Type", labels = c("Calcification", 
                                                 "Gross Photosynthesis", 
                                                 "Respiration"))
  
coral_plot %>%
  ggplot(aes(x=Activation_Energy, y=Rate)) +
  geom_point(aes(color=Site)) +
  facet_wrap(rate.type~., nrow=3, ncol=1, scales = "free") +
  scale_color_manual(values=unname(piratepal(palette = "pony", length.out = 6))) +
  theme_bw() 
  theme(legend.position = "bottom", plot.background = element_rect(fill=NULL), 
        strip.background = element_blank(), strip.text.x = element_blank(), 
        axis.text = element_blank(), axis.ticks = element_blank())
  geom_text(data=coral_plot_labels, aes(label=Type), hjust="middle")
```

Bar Plot

```{r}
plot_data <-sites %>% 
  filter(parameter.measured %in% c("temp", "light"))

plot_labels <- plot_data %>% group_by(name) %>% summarise(values=mean(values))
```

```{r}
plot_data %>% 
  ggplot(aes(x=name, y=values)) +
  geom_bar(stat = "identity", aes(color=parameter.measured), 
           fill = "cyan") +
  theme_void() + 
  scale_y_continuous(trans = "reverse") + 
  geom_text(data=plot_labels, aes(label=factor(name)), 
            position = position_dodge(width=4), 
            color = "orchid", size=9,
            fontface="bold", family="PartyLetPlain"
            ) +
  theme(plot.background = element_rect(fill="darkorchid"), 
        legend.position = "top", legend.title = element_blank(),
        legend.text = element_text(color = "cyan"), 
        legend.spacing.x = unit(3, "lines"),
        legend.box.background = element_rect(color="aquamarine", size=0.5, fill="darkorchid"))

```

```{r}
plot_data %>% 
  ggplot(aes(x=name, y=values, fill=parameter.measured)) +
  geom_bar(position="dodge", stat = "identity") + 
  facet_wrap(parameter.measured~., scales = "free_y", nrow=2, ncol=1) +
  theme_bw() +
  theme(legend.position = "none")
```
