---
title: "13a_Lab"
author: "Alyssa"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab

You have a set of 4 .csv files in data/homework. 
Each of these files is a timeseries of temperature and light data collected in tide pools in Oregon by Jenn Fields. 

Your goal is to 

-   bring in all 4 files
-   calculate the mean and standard deviation of both temperature (Temp.C) and light (Intensity.lux) for each tide pool.  
-   Use **both** a for loop and map() functions in your script. (Basically, do it twice). 
\
Due Monday at 9am. 


# Libraries
```{r}
library(tidyverse)
library(here)
```


# Data

**Data Dictionary**

|Variable Name | Description|
|:----------:|:------------|
|PoolID| ID of the pool|
|Foundation_spp| Surfgrass or mussel dominated pool|
|Removal_Control| Was it a removal or control treatment|
|Date.Time| Date and time|
|Temp.C| Temperature in degrees C|
|Intensity.lux| Light level in lux|
|LoggerDepth| Depth of the logger in meters|
\


# for loop

## List files
```{r}
# point to the location on the computer of the folder
tidepoolsPath<-here("Week_13", "Data", "lab")

# list all the files in that path with a specific pattern
files <- dir(path = tidepoolsPath, pattern = ".csv")
files
```

## pre-allocate space for the loop
```{r}
# pre-allocate space
# make an empty dataframe that has one row for each file and 3 columns
tidepools_data<-data.frame(matrix(nrow = length(files), ncol = 5))
# give the dataframe column names
colnames(tidepools_data)<-c("filename","mean_temp", "sd_temp", "mean_light", "sd_light")

tidepools_data
```

## calculate a mean and sd
```{r}
raw_data<-read_csv(paste0(tidepoolsPath,"/",files[1])) # test by reading in the first file and see if it works
head(raw_data)
mean(raw_data$Temp.C, na.rm = TRUE) # calculate a mean
sd(raw_data$Temp.C, na.rm = TRUE) # calculate sd

mean(raw_data$Intensity.lux, na.rm = TRUE)
sd(raw_data$Intensity.lux, na.rm = TRUE)
```

## make the loop
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 

raw_data<-read_csv(paste0(tidepoolsPath,"/",files[1]))
#glimpse(raw_data)

tidepools_data$filename[i]<-files[i]
tidepools_data$mean_temp[i]<-mean(raw_data$Temp.C, na.rm = TRUE)#<<
tidepools_data$sd_temp[i]<-sd(raw_data$Temp.C, na.rm = TRUE)#<<

tidepools_data$mean_light[i]<-mean(raw_data$Intensity.lux, na.rm = TRUE)#<< 
tidepools_data$sd_light[i]<-sd(raw_data$Intensity.lux, na.rm = TRUE)#<< 
} 

tidepools_data
```


# purrr

## files
list files
```{r}
files_p <- dir(path = tidepoolsPath,pattern = ".csv", full.names = TRUE)
files_p
```

## read in files and calculate means
```{r}
data <- files_p %>%
  set_names() %>%# set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map to dataframe with id in filename
  group_by(filename) %>% #<<
  summarise(mean_temp = mean(Temp.C, na.rm = TRUE), #<< add calculations
            sd_temp = sd(raw_data$Temp.C, na.rm = TRUE),
            mean_light = mean(Intensity.lux, na.rm = TRUE),
            sd_light = sd(Intensity.lux, na.rm = TRUE)) #<<

data
```




















