---
title: "12b_Lab"
author: "Alyssa"
date: "4/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


# Lab Outline

Working with intertidal data. Today you will need to use skills that you have learned in working with words to 

-   clean up some issues with the character data 
-   and use what you learned today to reorder your factors into orders that make sense.  
-   (Also the column headers could be better...)
\

**Goal:** 

-   make a plot
-   one of the axes is a factor.  
-   Save everything in the appropriate folders.

Data:
You have two possible files to play with in the data folder (you don't HAVE to use both): 

1.    intertidaldata.csv 
2.    intertidaldata_latitude.csv 

(as well as a data dictionary - intertidal_data_dictionary.csv)
\
**Importantly**- if you decide to visualize anything by _tide height_ it should go from low to mid to high tide.   
\
Due Monday, 9am.


# Libraries
```{r}
library(tidyverse)
library(here)
library(flextable)
library(ggridges)
```


# Data
```{r}
data_dictionary <- read_csv(here("Week_12", "Data", "intertidal_data_dictionary.csv"))
data_dictionary %>%
  flextable() %>% 
  theme_booktabs() %>%
  autofit()
```


```{r}
tides <- read_csv(here("Week_12", "Data", "intertidaldata.csv"))

lat <- read_csv(here("Week_12", "Data", "intertidaldata_latitude.csv"))
```


# clean up the data

only looking at the invertebrates 
```{r}
inverts <- tides %>% select(c(1:3, 11:13)) %>% 
  transform(Quadrat = str_replace_all(Quadrat, "\\.|[0-9]", "")) %>%
  mutate(Quadrat = str_trim(Quadrat), 
         Quadrat = factor(Quadrat, levels = c("Low", "Mid", "High"))) %>%
  rename("Whelks" = "Whelks..Counts.", 
         "Crabs" = "Crabs..Counts.", 
         "Sea Stars" = "Stars..Counts.") %>%
  pivot_longer(cols = c(4:6), 
               names_to = "Invertebrates", 
               values_to = "Count")

unique(inverts$Quadrat)  
```


# Plot
```{r}
inverts %>% 
  filter(Invertebrates == "Crabs", # only grabs
         Count != 0) %>% # remove 0s
  left_join(lat) %>%  # add in latitude to organize
  ggplot(aes(x=Count, 
             y=fct_reorder(Site, Latitude),
             fill=Invertebrates, 
             label= as.character(Latitude))) +
  geom_density_ridges(alpha=0.7) +
  scale_x_continuous(limits = c(0,5.5), 
                     expand = c(0,0)) + 
  labs(y="Site", title="Crabs Found at Each Site") +
  geom_text(aes(label=Latitude, x=5, y=Site), # add latitude labels
            size=2, face="plain", color="gray30") +
  coord_cartesian(clip = "off") + 
  annotate("text", label="Latitude", x=5.5, y=6.7, angle=-90, hjust=0.5) + # add title on right
  theme_classic() +
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(), 
        plot.margin = unit(c(1,2,1,1), "lines"), 
        plot.title.position = "plot") +
  ggsave(here("Week_12", "Outputs", "Crabs_12b.png"))
```




